---
title: "Plotting Vaccination Intention"
authors: "Zoey Dlott & Megan Richard""
subtitle: "An intro to R data viz in ggplot2"
output: html_document

---

## BEFORE YOU BEGIN

* Make sure you have the following packages installed: tidyverse, here
* Save this R-Markdown file AND the vaccine_intention.csv file in the SAME FOLDER on your computer (the code below will only load the data if it is saved in the folder *with* this document)

## Load needed packages

``` {r chunk1}
library(package = "here")
library(package = "tidyverse")
```

## Import the data set

``` {r chunk2}
# use here to get R to look in the same folder for the data
here()
# import the data
vaxxData <- read_csv("vaccine_intention.csv")
```

## Explore the data

```{r}
# change character variable types to categorical
vaxxData <- vaxxData %>%
  mutate_if(is.character, as.factor)
# show the data
summary(vaxxData)
```

## Basic Bar Graph

```{r}
vaxxData %>%
  ggplot(aes(x = vaccination.intention)) +
  geom_bar() 
```

## Fancier Bar Graph

```{r}

vaxxData %>%
  ggplot(aes(x = vaccination.intention)) +
  geom_bar(fill = "slategray1") + 
  labs(x = "Vaccination Intention ",
       y = "Number of People",
       title = "Figure 1: COVID-19 Vaccination Intention") +
  theme_minimal()
```

# You Try it!

Make a bar graph of **flu shot history** with the Figure 1 code in your R Markdown file.

Edit the code to: 

* Change the color of the bars by replacing "slategray1" the name of your favorite color (e.g., orange) or using a color found on the https://colorbrewer2.org/ website

* Add labels for the axes and a title inside the quote marks

* Try another theme, you can find them here: https://ggplot2.tidyverse.org/reference/ggtheme.html

When you have completed the changes, click the little green arrow on the top right of the code chunk to run the code.

```{r}
# Figure 1
vaxxData %>%
  ggplot(aes(x = flu.shot.history)) +
  geom_bar(fill = "slategray1") + 
  labs(x = "", y = "", title = "") +
  theme_minimal()

```

## Basic Box Plot

```{r}

vaxxData %>%
  ggplot(aes(x = vaccination.intention, y = age)) +
  geom_boxplot()

```

## Adding Layers to Box Plot
```{r}

vaxxData %>%
  ggplot(aes(x = vaccination.intention, y = age, 
             fill = vaccination.intention)) +
  geom_boxplot() +
  labs(x = "Vaccination Intention ",
       y = "Age",
       title = "Vaccination intention by Age") +
  theme(legend.position = "none")

```

## Pretty But Not Useful Bar Plot

```{r}

vaxxData %>%
  ggplot(aes(x = vaccination.intention, 
             fill = flu.shot.history)) +
  geom_bar() +
  scale_fill_brewer(palette = 'Set3', 
                    name = "Flu Shot History") +
  labs(x = "Vaccine Intention", 
       y = "Number of Individuals",
       title = "Figure 2: Relationship between COVID\nVaccine Intention and Flu Shot History") +
  theme_minimal()


```

## Basic Grouped Bar Plot

```{r}

vaxxData %>%
  group_by(flu.shot.history, vaccination.intention) %>%
  count() %>%
  group_by(vaccination.intention) %>%
  mutate(perc = 100*n/sum(n)) %>%
  ggplot(aes(x = vaccination.intention, y = perc, 
             fill = flu.shot.history)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(x = "Vaccine Intention", 
       y = "Percent of Individuals",
       title = "Figure 3: Relationship between COVID\nVaccine Intention and Flu Shot History") +
  scale_fill_manual(values = c("skyblue", "cornflowerblue"),
                    name = "Flu Shot History")


```

# You Try it!

Customize the bar graph of the vaccination intention and flu shot history with the Figure 2 code in your R Markdown file.

* Copy and paste the code for flu.shot.history and Edit the code to:

* Change the flu shot history variable to the `political.beliefs` variable

* Change the colors of the bars by replacing "plum2", "plum3", "plum4" with the names of your favorite colors (e.g., orange) and add 3 more colors for all the categories of political beliefs. Try finding colors on the https://colorbrewer2.org/ website (HINT: Political beliefs has 6 colors, you'll need to add 3 more colors to the list!)

* Add labels for the axes and a title inside the quote marks

* Try another theme, you can find them here: https://ggplot2.tidyverse.org/reference/ggtheme.html

When you have completed the changes, click the little green arrow on the top right of the code chunk to run the code.

```{r}

```

## Answer 2 Clarify

```{r}

vaxxData %>%
  group_by(political.beliefs, vaccination.intention) %>%
  count() %>%
  group_by(political.beliefs) %>%
  mutate(perc = 100*n/sum(n)) %>%
  mutate(political.beliefs = factor(political.beliefs, 
                                    levels=c("Very Liberal", 
                                             "Somewhat Liberal", 
                                             "Mixed", 
                                             "Somewhat Conservative",  
                                             "Very Conservative",
                                             "Prefer not to answer"))) %>%
  ggplot(aes(x = political.beliefs, y = perc, 
             fill = vaccination.intention)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(x = "Political Beliefs", 
       y = "Percent of Individuals",
       title = "Relationship between COVID Vaccine Intention\nand Political Beliefs") +
  scale_fill_manual(values = c("plum2","plum3","plum4"),
                    name = "Vaccination Intention",
                    guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  coord_flip()

```


## Basic Histogram

```{r}

vaxxData %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 1.5) +
  scale_fill_brewer(palette = 'Set2', 
                    name = "Ages of Participants") 

```

## Fancy Histogram

```{r my_1var_histo_fancy, include = FALSE}

vaxxData %>%
  ggplot(aes(x= age)) + 
  geom_histogram(breaks=seq(10, 60, by = 1), 
                 col="blue", 
                 fill="slategrey", 
                 alpha = .2) + 
  labs(title="Histogram of Age") +
  labs(x="Age", y="Count") + 
  xlim(c(18, 60)) 

```

## Basic Violin Plot

```{r}

vaxxData %>% 
  ggplot(aes(x = vaccination.intention, y = age)) +
  geom_violin(aes(fill = vaccination.intention)) +
  labs(y = "Age in Years", 
       x = "Do You Intend to get the COVID-19 Vaccine?",
       title = "Figure 7: Vaccination Intention by Age") + 
  scale_fill_brewer(palette = "Set2", guide = FALSE) + 
  theme_minimal()

```

## Horizontal Violin Plot

```{r}

vaxxData %>% 
  ggplot(aes(x = political.beliefs, y = age)) +
  geom_violin(aes(fill = political.beliefs)) +
  labs(y = "Age in Years", x = "Political Beliefs",
       title = "Figure 9: Political Beliefs by Age") + 
  scale_fill_brewer(palette = "Set2", guide = FALSE) + 
  coord_flip() +
  theme_minimal()

```

## Super Fancy Boxplot & Violin Plot

```{r}

vaxxData %>%
  ggplot(aes(x = vaccination.intention, y = age)) +
  geom_violin(aes(color = vaccination.intention), fill = "white",
              alpha = .8) +
  geom_boxplot(aes(color = vaccination.intention, fill = vaccination.intention), 
               alpha = .3) +
  geom_jitter(aes(color = vaccination.intention), alpha = .4) +
  labs(x = "Vaccination Intention",
       y = "Age (in years)",
       title = "Figure 6: Vaccination Intention by Age") +
  scale_color_brewer(palette = "Set2", guide = FALSE) +
  scale_fill_brewer(palette = "Set2", guide = FALSE) +
  theme_minimal(base_size = 14)

```

# You Try it!

Edit the Super Fancy Boxplot & Violin Plot code to: 

* Change the categorical variable (keep age). Remember to change it everywhere it shows up!

* Change the colors by replacing the Set2 palette with a different palette found on the http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/ website (bottom of page). Remember to change it everywhere it shows up!

* Add labels for the axes and a title inside the quote marks

* Try another theme, you can find them here: https://ggplot2.tidyverse.org/reference/ggtheme.html

When you have completed the changes, click the little green arrow on the top right of the code chunk to run the code.

```{r}

vaxxData %>%
  ggplot(aes(x = vaccination.intention, y = age)) +
  geom_violin(aes(color = vaccination.intention), fill = "white",
              alpha = .8) +
  geom_boxplot(aes(color = vaccination.intention, fill = vaccination.intention), 
               alpha = .3) +
  geom_jitter(aes(color = vaccination.intention), alpha = .4) +
  labs(x = "Vaccination Intention",
       y = "Age (in years)",
       title = "Figure 6: Vaccination Intention by Age") +
  scale_color_brewer(palette = "Set2", guide = FALSE) +
  scale_fill_brewer(palette = "Set2", guide = FALSE) +
  theme_minimal(base_size = 14)

```

## The End!